---
title: "Chris_Almeria_assignment02"
author: "Chris Almeria"
date: "9/3/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# Direct upload to Github
# Analyzing Data: Conditional Means for home ownership rates

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(Metrics)
```

## Dataset of U.S. Counties 

```{r data}
load("pd.Rdata")
```

## Full explination of data parameters for this data set

```{r explain_date}

## Full explanation of data in codebook
load("pd_lab_explain.Rdata")

#or use View
View(lab_explain)
```

## Rank order counties by home ownership rates for easier visual repesentation. 
```{r simple_plots}
## Creating a rank variable for home Ownership rates, by way of the mutate command
pd<-pd%>%mutate(homeown_rate=rank(homeown_rate))
```

```{r}
## Plot by rank
gg<-ggplot(data=pd, aes(x=homeown_rate,
                         y=median_hh_inc,
                         text=county))
```

# Scatterplot
```{r}
## Add Points
gg<-gg+geom_point(alpha=.5,size=.5)
```

# Lables
```{r}
## Add labels
gg<-gg+xlab("Rank")+ylab("Median Household Income")
```


# Naming Convention, gg
```{r}
gg
```
# Making this plot interactive to mouse over and view college grad rates and homeownership at the same time by way of the ggplotly command.

```{r}
# Make Interactive plot
gg_p<-ggplotly(gg)

gg_p
```

## Dependent Variable: Home Ownership Rates (homeown_rate)

```{r}

## Question 1
## Calculate mean of the outcome, Home Ownership Rates by way of the summarize pipe function/summarize command
pd%>%summarize(mean_homeown_rate=mean(homeown_rate,na.rm=TRUE))

## Question 2
## Using this mean from question one as a predictor, by way of the mutate command.
pd<-pd%>%mutate(mean_homeown_rate=mean(homeown_rate,na.rm=TRUE))
```
## Plotting the above commands visually, with Homeownership rates on the x axis, College graduation rates on the Y axis
## Made a mistake here by mutating this command to early as the median value for Homeownership Nation wide is the only one that shows (1544.5)?

```{r}
##Plotting
gg<-ggplot(data=pd,aes(y=median_hh_inc,x=homeown_rate,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(y=median_hh_inc,x=homeown_rate,
                  color="Predicted: Unconditional Mean"),
                  size=.5)
gg<-gg+xlab("Homeownership Rate Median")+ylab("Median Household Income")
gg<-gg+scale_color_manual(name="Type",
                          values=c("Actual"="black",
                          "Predicted: Unconditional Mean"="blue")
                          )
gg<-gg+theme(legend.position="bottom")

gg

##Save for later

gg2<-gg

```
## Question 3
## Calculating a summary measure of the error of the home ownership rate figure and storing it in the value e1
```{r error_terms}
pd<-pd%>%mutate(e1=homeown_rate-mean_homeown_rate)
```
## Question 4
## Calculating the mean of our predictor variable, staying with median household incomes
## Using the Root Mean Squared Error function to see how far 'off' this measure is: 891.4178
```{r}
## RMSE

rmse_uncond_mean<-rmse(pd$homeown_rate,pd$mean_homeown_rate)

rmse_uncond_mean
```
## Question 5
## Using these Condtional means as a predictor 
## Adapted code to use the median household income level as a basis to look at homeownership rates - thought being the higher HH income rates, we should expect to see a higher rate of homeownership.
```{r condtl_mean_single}
##Condtional Average across a single variable

## Creating a new variable for quartiles of median household incomes by wy of the mutate command
pd<-pd%>%mutate(median_hh_level=ntile(median_hh_inc,4))

pd%>%select(county,median_hh_inc,median_hh_level)%>%View()

table(pd$median_hh_level)

##pd<-pd%>%mutate(median_hh_level=ntile(median_hh_inc,4))

pd<-pd%>%group_by(median_hh_level)%>% ## Group by predictor
  ##Calculate mean at each level of predictor
  mutate(pred_median_hh_inc=mean(percapinc.2010))%>%
  ## Ungroup
  ungroup()%>% 
  #Rank by prediction, with ties sorted randomly
  mutate(pred_median_hh_inc_rank=rank(pred_median_hh_inc,ties.method="random"))

pd%>%select(county,median_hh_inc,median_hh_level,pred_median_hh_inc)%>%View()
 
```
## Graphing these results out:
```{r}
pd%>%group_by(median_hh_level)%>% ## Group by predictor
  ##Calculate mean at each level of predictor
  summarise(pred_median_hh_inc=mean(percapinc.2010))

gg<-ggplot(data=pd,aes(x=pred_median_hh_inc,y=percapinc.2010,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(x=pred_median_hh_inc_rank,y=pred_median_hh_inc,color="Predicted:Conditional Mean, 1 var"))
gg<-gg+ scale_color_manual("Type",values=c("Predicted:Conditional Mean, 1 var"="red","Actual"="black"))
gg<-gg+theme(legend.position="bottom")
gg<-gg+xlab("Rank")+ylab("Per Capita Income, 2010")
gg

##Save for later
gg3<-gg

```
# Question 6
## Updating RMSE now that we are using an conditional vs an uncondtional mean and a summary of the error in these predictions. 
```{r}
rmse_cond_mean_one<-rmse(pd$median_hh_level,pd$pred_median_hh_inc)
rmse_cond_mean_one
```

## Overall, once I split the median income into quadrants, the outliers with high median incomes (King County, WA - Seattle & Bellvue; Los Angeles County, CA - LA and Hollywood, were 'constratined' by the upper quartiles max of $40,990, but such is to be expected when hanressing such a vast array of numbers)
## I chose to go with median household income on the basis of a higher income would allow people to own thier won homes vs rent, as this would also dovetial with college grad rates, since higher wages are typically earned by college grads
